# Understanding Network Data Structures

**Learning objectives:**

-   Generating networks from edge lists or adjacency matrices.

## Edge lists

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(igraph))
```

-   The list of relationships underlying a network among it's components is called **edge lists**.
-   An edge is defined by the pair of agents linked together. In other words, for each edge, **two** pieces of information are needed.
-   An edge list for a network can be defined by a table of two columns where the number of rows equal the number of edges.
-   In a directed graph, the nodes in first column are the sources of edges, and nodes in the second column receives the link.
-   In an undirected graph, order of columns doesn't matter.

```{r}
personA <- c("Mark", "Mark", "Peter", "Peter", "Bob", "Jill")
personB <- c("Peter", "Jill", "Bob", "Aaron", "Jill", "Aaron")

edgelist <- data.frame(PersonA = personA, PersonB = personB, stringsAsFactors = F)

print(edgelist)
```

Pros:

-    Simple and intuitive. - Number of rows equal the number of edges.

Cons:

-   Impossible to represent isolates using an edge list since it details relations.[^-1]

-   Doesn't support analysis methods.

[^-1]: Pierre explained the meaning of "isolates" which are unlinked nodes.

## Adjacency matrices (recommended structure)

```{r}
adjacency <- matrix(c(0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0), nrow = 5, ncol = 5, dimnames = list(c("Mark", "Peter", "Bob", "Jill", "Aaron"), c("Mark", "Peter", "Bob", "Jill", "Aaron")))

print(adjacency)
```

Pros:

-   More efficient than edge lists (Example: searching connections).

Cons:

-   More suited form computers than human (Example: difficult to record).

Convert edge list to adjacency matrix

```{r}
edgelist %>% 
  as.matrix() %>% 
  graph_from_edgelist(directed = FALSE) %>%
  as_adjacency_matrix(sparse = TRUE)
```

## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>

<summary>

Meeting chat log

</summary>

    LOG

</details>
