# Your First Network

## Creating a project

Show a quick demo.

## Loading data into R

For reading large datasets, use one of the options below

[vroom](https://www.tidyverse.org/blog/2019/05/vroom-1-0-0/)

[fread](https://rdatatable.gitlab.io/data.table/reference/fread.html)

## Manual data entry

-   Practically you'll never need, or advised, to enter values manually.
-   You can edit dataframes interactively within R using one of the following options

```{r, eval = FALSE}
fix(mtcars)
```

```{r, eval = FALSE}
DataEditR::data_edit(mtcars)
```

## From data to networks

```{r}
suppressPackageStartupMessages(library(igraph))
```

```{r}
#edge list
personA <- c("Mark", "Mark", "Peter", "Peter", "Bob", "Jill")
personB <- c("Peter", "Jill", "Bob", "Aaron", "Jill", "Aaron")

edgelist <- cbind(PersonA = personA, PersonB = personB)

print(edgelist)

#adjacency matrix
adjacency <- matrix(c(0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0),
                    nrow = 5,
                    ncol = 5,
                    dimnames = list(c("Mark", "Peter", "Bob", "Jill", "Aaron"),
                                    c("Mark", "Peter", "Bob", "Jill", "Aaron")))

print(adjacency)
```

Edge list into a network object.

```{r}
network <- graph.edgelist(edgelist, directed=FALSE)
network
```

Adjacency matrix to network object.

```{r}
network_adj <-graph.adjacency(adjacency,mode = "undirected" )
```

-   `UN`: undirected
-   `DN` : directed
-   #vertices #edges

Vertices of network

```{r}
V(network)
```

Edges of network

```{r}
E(network)
```

We can visualize the network using the plot() function.

```{r}
plot(network)
```
