---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Calculating Network Size and Density

**Learning objectives:**

- Measure graph density.

Load libraries
```{r message=FALSE}
library(igraph)
library(tidyverse)
```

## Load data

Regenerate data from previous chapter. (code below copied from previous chapter)
```{r}
#read data
gss_url <-"https://raw.githubusercontent.com/mahoffman/stanford_networks/main/data/gss_local_nets.csv"
gss <- read_csv(gss_url) 

#subset ties data
ties <- gss %>%
  dplyr::select(starts_with("close"))

#set zero links as missing links
ties[ties==0] <- NA
#repondents where all links among others are missing
others_missing <- rowSums(is.na(ties))==ncol(ties)
#remove any respondent that falls in any of the above
ties <- ties[!(others_missing),]

#make a function to generate a network of ties
make_ego_nets_simple <- function(tie){
  #get the all possible links among others
  tie <- tie %>% unlist
  #remove missing links
  tie <- tie[!is.na(tie)]
  #remove zero links
  tie <- tie[tie!=0]
  #get the identity of linked pairs
  others <- str_extract(names(tie), "[0-9]+")
  #split the linked others
  others_link <- str_split(others, "",simplify = TRUE) %>% as.data.frame
  #make edge list of others
  others_link <- cbind(others_link, tie)
  #ego graph with 
  graph_from_data_frame(others_link,
                        directed=FALSE)
}
#get all graphs
ego_nets_simple <- apply(ties,1,make_ego_nets_simple)
#
ties[1,]
plot(ego_nets_simple[[1]])

```

## Number of vertices

Number of vertices
```{r}
network_sizes <- lapply(ego_nets_simple, vcount)
head(network_sizes)
```

```{r}
network_sizes <- unlist(network_sizes)
mean(network_sizes, na.rm = T)
```

```{r}
hist(network_sizes, main = "Histogram of Ego Network Sizes", xlab = "Network Size")

```

## Number of edges

Number of edges
```{r}
network_edge_counts <- lapply(ego_nets_simple, ecount)
network_edge_counts <- unlist(network_edge_counts)
hist(network_edge_counts, main = "Histogram of Ego Network Edge Counts", xlab = "# of Edges")
```

```{r}
random_ego_net <- ego_nets_simple[[1]]
ecount(random_ego_net)/((vcount(random_ego_net) * (vcount(random_ego_net) - 1))/2)

```

## Density 

From the documentation of `graph.density()` "The density of a graph is the ratio of the number of edges and the number of possible edges."
```{r}
densities <- lapply(ego_nets_simple, graph.density)
densities <- unlist(densities)
```

```{r}
hist(densities)
```

## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
