# Affiliation Data

**Learning objectives:**

-   THESE ARE NICE TO HAVE BUT NOT ABSOLUTELY NECESSARY

```{r message=TRUE, warning=TRUE}
library(tidyverse)
library(tidygraph)
library(ggraph)
library(igraph)
library(nberwp)
library(nberwp)
```

## Indirect connections

Linking individuals indirectly based on shared features.

Example: Affiliation data.

```{r}
name <- c("Leo", "Clement", "Palla", "Filippo")
class1 <- c("Biostatistics","Islamic Civ", "Calc 1", "Linear Algebra")
class2 <- c("Chemistry", "The Modern World-System", "Calc 2", "Social Networks")
class3 <- c("Linear Algebra", "Exile and Diaspora", "Linear Algebra", "The Modern World-System")
(classes_data <- bind_cols(name = name,
                           class1 = class1,
                           class2 = class2,
                           class3 = class3))
```

Get edgelist by converting from wide to long format

```{r}
(classes_data <- classes_data %>%
  pivot_longer(-name,
               names_to = NULL, #discard colum names
               values_to = "classes"))
```

Convert edgelist into an incidence matrix.

```{r}
(classesMatrix <- classes_data %>% 
  sjmisc::flat_table() %>% 
  as.matrix())
```

In an incidence matrix, the rows are of one class of node, while columns are of another. The rows are generally people who are affiliated with groups in the columns.

Using the get.incidence() function will turn our matrix into a bipartite network.

```{r}
(classesNet_tg <- graph.incidence(classesMatrix,
                                 mode = c("all")#all mutual edges 
                                 ) %>% 
  as_tbl_graph())
```

```{r}
(classesNet_tg <- classesNet_tg %>% 
  mutate(type = ifelse(type, "Affiliations", "Individuals")))
```

```{r}
ggraph(classesNet_tg,  layout = 'kk') + 
  geom_edge_link() + 
  geom_node_label(aes(label = name, fill = type))+
  guides(fill = guide_legend(
    override.aes = aes(size = 5,
                       label = "")
    ))+
  theme(legend.position = c(0.1,0.1),
        legend.text = element_text(size = 10))
```


### Unipartite Projection

Using igraph `bipartite.projection`

```{r}
graph.incidence(classesMatrix,
                mode = c("all")#all mutual edges 
                ) %>% 
  bipartite.projection(which = "false") #select the projection of interest. Here it's students
```

The mathematical operation to make a person-to-person projection is to multiply the initial matrix by its transpose. In R that looks like:
```{r}
personMatrix <- classesMatrix %*% t(classesMatrix)
```


The diagonal of this new matrix tells us the number of groups each person is affiliated with, but we set it to 0 using the ?diag function.

```{r}
number_of_classes_taken = diag(personMatrix)
diag(personMatrix) <- 0 
```

```{r}
personMatrix %>%
  graph.adjacency() %>% 
  as_tbl_graph() %>% 
  ggraph( layout = 'kk') + 
  geom_edge_link() + 
  geom_node_label(aes(label = name))
```

To get the group-to-group matrix, we multiply the transpose by the initial ma

trix (reverse!)

```{r}
groupMatrix = t(classesMatrix) %*% classesMatrix
number_of_students <- diag(groupMatrix)
diag(groupMatrix) <- 0 # we again set it to 0
```

```{r}

groupMatrix %>%
  graph.adjacency() %>% 
  as_tbl_graph() %>% 
  ggraph( layout = 'kk') + 
  geom_edge_link() + 
  geom_node_label(aes(label = name))
```

## Tripartite network analysis?

How to add the respective department of each class to the story?

```{r}
classes_to_departments <- data.frame(class = c("Biostatistics","Islamic Civ", "Calc 1", "Linear Algebra", "Chemistry", "The Modern World-System", "Calc 2", "Social Networks", "Exile and Diaspora"), department = c("Math", "History", "Math", "Math", "Chemistry", "Sociology", "Math", "Sociology", "History"), stringsAsFactors = F)
classes_to_departments_matrix <- table(classes_to_departments)
class(classes_to_departments_matrix) <- "matrix"  
```

Now following the paper on tripartite structural analysis, we can multiply the transpose of this matrix by the transpose of classesMatrix to trace links between people and departments!

```{r}
people_to_departments <- t(classes_to_departments_matrix) %*% t(classesMatrix)
```

We can graph this matrix and analyze it like a bipartite graph.

```{r}
people_to_departments %>%
  graph.incidence() %>% 
  as_tbl_graph() %>% 
  ggraph( layout = 'kk') + 
  geom_edge_link() + 
  geom_node_label(aes(label = name, fill = type))
```

### Lab

```{r}
data(paper_authors)
data(authors)
data(papers)
papers <- papers %>%
  filter(year >= 2016)
paper_authors <- paper_authors %>% 
  filter(paper %in% papers$paper)
paper_authors <- paper_authors %>% 
  mutate(author = authors$user_nber[match(paper_authors$author , authors$author)])
```

```{r eval = FALSE}
paper_authors %>% 
  sjmisc::flat_table() %>% 
  graph.incidence() %>% 
  as_tbl_graph() %>% 
  ggraph( layout = 'kk') + 
  geom_edge_link() + 
  geom_node_poin()
```

## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/wU-g5vaXgCU")`
